---
title: "Druglogics ROC curve Generator"
author: "[John Zobolas](https://github.com/bblodfon)"
output:
  html_document:
    df_print: paged
    theme: united
    toc: true
    toc_float: true
runtime: shiny
---

# Intro

With this code we can **generate a ROC curve** from one or two ensemble-wise 
synergy files, which is one of the output results of the druglogics pipeline 
module `drabme`.

# Prerequisites

```{r Load useful functions, message = FALSE}
source("useful_fun.R") # loads `dplyr`
```

# Example Input

**Input files** that are needed in order to generate the ROC curve are:

- One **observed synergies file** to test against the synergies found from the 
`drabme` simulation(s)
- At least **one ensemble-wise synergy** result file
  - In case of one such file, the ROC curve can be drawn using either the 
  **original** synergy score results or the **random** ones
  - In case of two such files, the second (**random**) is used for normalizing 
  the results of the first (**original**) through various methods:
    - **exp-prod-norm**: $orig\cdot rand\Rightarrow exp(orig)\cdot exp(rand)=exp(orig+rand)$
    - **exp-fold-change-norm**: $orig/ rand\Rightarrow exp(orig)/ exp(rand)=exp(orig-rand)$

Check in the `input/` dir for example files and their format.  
In the next section we show how to create a combined `predictions` object from
such files:
```{r Input and validation}
input.dir = paste0(getwd(), '/input')
original.file = paste0(input.dir, "/ensemble_synergies")
random.file = paste0(input.dir, "/ensemble_synergies_random")
observed.synergies.file = paste0(input.dir, "/observed_synergies")

# load files
observed.synergies = get.observed.synergies(observed.synergies.file)
orig.res = read.ensemble.synergies.drame.file(original.file)
rand.res = read.ensemble.synergies.drame.file(random.file)

data.validation(orig.res, rand.res, observed.synergies)
predictions.example = 
  create.predictions.data.frame(orig.res, rand.res, observed.synergies)
```

You can of course use the rest of the code if you can create a `predictions` 
object yourself (with at least one of the `original` or `random` columns).  

Get a glimpse of the `predictions.example` data:
```{r Example predictions (embeded Sniny input and ouput), echo = FALSE}
numericInput(inputId = "rows", label = "How many rows to show?", value = 5,
             min = 0, max = nrow(predictions.example))

renderTable({
  head(predictions.example, n = input$rows)
}, digits = 7)
```

# Create ROC Curve

Choose a single or a normalization method and input files (at least one drabme 
file is needed, otherwise the example input from above is used) to draw the ROC 
curve:
```{r Sniny App To Be, echo = FALSE}
# Input section
selectInput(inputId = "method", label = "Select Method", 
            choices = c("original", "random", "exp-prod-norm", 
                        "exp-fold-change-norm"))

fileInput(inputId = "observed", label = "Select Observed Synergies File", 
          buttonLabel = "Browse...", placeholder = "No file selected")
fileInput(inputId = "orig", label = "Select Original Ensemble Synergies File", 
          buttonLabel = "Browse...", placeholder = "No file selected")
fileInput(inputId = "rand", label = "Select Random Ensemble Synergies File", 
          buttonLabel = "Browse...", placeholder = "No file selected")

# drawing section
renderPlot({
  # take input
  method = input$method
  observed.synergies.file = input$observed$datapath
  original.file = input$orig$datapath
  random.file = input$rand$datapath
  
  # which `predictions` object to use?
  if (!is.null(observed.synergies.file) & (!is.null(original.file) | 
                                           !is.null(random.file))) {
    # we generate new `predictions`
    observed.synergies = get.observed.synergies(observed.synergies.file)
    orig.res = read.ensemble.synergies.drame.file(original.file)
    rand.res = read.ensemble.synergies.drame.file(random.file)
    
    data.validation(orig.res, rand.res, observed.synergies)
    predictions = create.predictions.data.frame(orig.res, rand.res, observed.synergies)
  }
  else {
    # use example `predictions`
    predictions = predictions.example
  }
  
  # validate if input method is combatible with `predictions` data
  method.check(method, predictions)
  method.column = map.method.to.column(method)
  
  # normalize
  predictions = normalize(predictions)
  
  # generate ROC stats and curve
  roc_stats = gen.roc.stats(predictions, method.column)
  x = roc_stats$FPR
  y = roc_stats$TPR
  AUC = sum(diff(x) * (head(y,-1)+tail(y,-1)))/2
  
  plot(x, y, type = "b", col = "blue", 
       main = "ROC curve", xlab = "False Positive Rate (FPR)", 
       ylab = "True Positive Rate (TPR)")
  legend("bottomright", legend = specify.decimal(AUC, digits.to.keep = 3), 
         title = "AUC", col = "blue", pch = 19)
  grid()
  abline(a = 0, b = 1, col = "lightgray", lty = 2) # y=bx+a
  
}, height = 500, width = 500)
```